### 2024-12-17 article/main.tex

%% josis.tex 1.4   2016-09-15    JoSIS latex template
%------------------------------------------------------------------
% Filename: josis_template.tex
%
% This file is intended as a template for typesetting articles for the
%
%                        Journal of Spatial Information Science.
%
% Please edit this template to generate your own formatted manuscripts
% for submission to JOSIS. See http://josis.org for further details.
%

%%% JOSIS checks in typesetting
%%% * All titles and sections lower case *EXCEPT short title  [ ]
%%% * Remove author postal addresses, only have geographic places and institutions [ ] 
%%% * Consistent use of Section, Figure, Table (capitalized and in full) [ ]
%%% * 10 keywords (and all lower case) [ ]
%%% * Remove all avoidable footnotes [ ]
%%% * Use double quotation marks (``'' not "" or `') [ ]
%%% * Punctuation inside quotations [ ]
%%% * E.g. and i.e. followed by comma [ ]
%%% * cf. followed by tilde [ ]
%%% * Itemize and enumerate correctly punctuated [e.g., "1. x, 2. y, and 3. x." ]
%%% * And/or lists using American English punctuation (e.g., "x, y, and z") [ ] 
%%% * Bibliography (e.g., en-dashes for number ranges, consistent "Proc.~" for Proceedings of..., etc.) []
%%% * Acknowledgment style use section* [ ] 
%%% * et al. no italics, but with dot  [ ] 
%%% * All captions end with full stop  [ ] 
%%% * Table captions under, not over table  [ ]
%%% * Adjust urls with burlalt [ ] 
%%% * Check correct use of hyphens, emdashes, endashes  [ ]
%%% * Perform spell check  [ ]

%%% JOSIS checks directly before publication 
%%% Check DOI, page numbers on article and web site. [ ]
%%% Update web site with final title, abstract, keywords. [ ] 
%%% Build with distiller for DOI links. [ ]

% Required documentclass definition for JOSIS
\documentclass{josis}
\usepackage{hyperref}
\usepackage[hyphenbreaks]{breakurl}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage[T1]{fontenc}
\usepackage{cite}

% Suggested packages for algorithm formatting
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}

\usepackage[table]{xcolor}
\usepackage{amssymb,amsmath}

\renewcommand{\topfraction}{0.9} 
\renewcommand{\textfraction}{0.1}

% Page setup and overhangs
\sloppy
\widowpenalty=10000
\clubpenalty=10000
\hyphenpenalty=75

% Article details for accepted manuscripts will be added by editorial staff
% Omit year if article in press
% Omit number if article under review
\josisdetails{%
   number=N, year=YYYY, firstpage=xx, lastpage=yy, 
  doi={10.5311/JOSIS.YYYY.II.NNN},
   received={December 18, 2024}, 
   returned={},
   revised={},
   accepted={}, }

\newcommand{\mydoi}[1]{\href{http://dx.doi.org/#1}{doi:\protect\detokenize{#1}}}

%\renewcommand{\UrlLeft}{http:\sslash}
%\DeclareUrlCommand\myurl{\def\UrlLeft{}\def\UrlRight{}%
%\urlstyle{tt}}

\urlstyle{rm}
\makeatletter
% Inspired by http://anti.teamidiot.de/nei/2009/09/latex_url_slash_spacingkerning/
% but slightly less kern and shorter underscore
\let\UrlSpecialsOld\UrlSpecials
\def\UrlSpecials{\UrlSpecialsOld\do\/{\Url@slash}\do\_{\Url@underscore}}%
\def\Url@slash{\@ifnextchar/{\kern-.11em\mathchar47\kern-.2em}%
    {\kern-.0em\mathchar47\kern-.08em\penalty\UrlBigBreakPenalty}}
\def\Url@underscore{\nfss@text{\leavevmode \kern.06em\vbox{\hrule\@width.3em}}}
\makeatother

\hypersetup{
colorlinks=true,
linkcolor=black,
citecolor=black,
urlcolor=black
}

% Add the running author and running title information
\runningauthor{\begin{minipage}{.9\textwidth}\centering Julian Kraft\end{minipage}}
\runningtitle{Wildlife sightings WebApp}

% Document begins
\begin{document}
%\setcounter{page}{33}

% Insert your own title
\title{A Web Application for Wildlife Sightings - how accessible is this technology using modern frameworks?}

% Insert your manuscipts authors, affiliations, and addresses
\author{Julian Kraft}\affil{LSFM, ZHAW, Switzerland}

\maketitle

% Add 5-10 keywords for every submission
\keywords{Web Application, Wildlife Sightings, Geodata, Open Source, Web Frameworks}

% Add a short abstract of 150-250 words
\begin{abstract}
The aim of this semester project is to assess the accessability of the technology to create a web application to record geo tagged data.
With the help of modern web frameworks, a web application was built to record wildlife sightings. A working prototype was created and and made
available online with the option to install it on smartphones of all major operating systems.
The project was a success and the technology is accessible to everyone with a bit more than basic programming knowledge and maybe some help.

\end{abstract}

% Your main text begins here. 
\section{Introduction}

\section{Author guidelines}

\section*{Acknowledgments}

Acknowledgments appear in a separate unnumbered section before the bibliography. Please acknowledge anyone (individual/company/institution) who has contributed to the study, including substantial contributions to the conception, design, acquisition of data; analysis and interpretation of data; drafting the manuscript; or provided critical comments resulting in revisions to content. For each author, please list the source(s) of any funding or financial contributions related to the study.

\bibliographystyle{josisacm}
\bibliography{josisexample}

\end{document}